---
title: Vidigi vs BupaR
execute:
  echo: true
filters:
   - include-code-files
engine: knitr
format:
  html:
    toc: true
    theme:
      light: [flatly, ../theme-light.scss]
      dark: [darkly, ../theme-dark.scss]
---

```{=html}
<style>
.processmapR {
  background-color: white;
  border-radius: 4px;
}

.processanimateR {
  background-color: white;
  border-radius: 4px;
}

.dataTables_wrapper {
  background-color: white;
  border-radius: 4px;
}

</style>
```

The title of this section is perhaps misleading! As the author of the package, I think the visuals produced by the two packages occupy slightly different niches, and the use of both can benefit your project.

As an additional bonus, the process of creating the logs you require for a vidigi project give you the perfect dataset for your bupaR visuals too!

bupaR outputs could form part of a verification and validation strategy. They can also perform part of your communications strategy, helping to provide a talking point for meetings with stakeholders in much the same way as a screenshot of your Simul8 or Anylogic model would. In the absence of a graphical interface for building a model, the bupar outputs can help you - and your stakeholders - to ensure that linkages between different model steps are sensible and appropriate.

## The Python Preprocesing

We will begin in Python, working to add a couple of columns to our vidigi event log to prepare it for use in bupaR.

```{python}
#| echo: true

from examples.example_2_branching_multistep.ex_2_model_classes import Trial, g
import pandas as pd
import os

#| echo: true

g.sim_duration = 3000
g.number_of_runs = 1
```

:::{.callout-note collapse="true"}
### View Imported Code, which has had logging steps added at the appropriate points in the 'model' class

```{.python include="../examples/example_2_branching_multistep/ex_2_model_classes.py"}
```

:::

```{python}
#| echo: true

my_trial = Trial()

my_trial.run_trial()

my_trial.all_event_logs.head(10)
```

```{python}
my_trial.all_event_logs.event_type.value_counts()
```

```{python}
my_trial.all_event_logs.event.value_counts()
```

```{python}
#| echo: true

# First, identify all patients who have a 'depart' event
# patients_with_depart = my_trial.all_event_logs[my_trial.all_event_logs['event'].str.contains('depart')]['patient'].unique()

# Then filter the original DataFrame to only include those patients
# filtered_df = my_trial.all_event_logs[my_trial.all_event_logs['patient'].isin(patients_with_depart)]

logs_transformed = my_trial.all_event_logs[~my_trial.all_event_logs['event'].str.contains('wait')].copy()
# logs_transformed = filtered_df[~filtered_df['event'].str.contains('wait')].copy()
logs_transformed = logs_transformed[logs_transformed['event_type'].isin(['resource_use', 'resource_use_end'])].copy()
logs_transformed['event_stage'] = logs_transformed['event_type'].apply(lambda x: 'complete' if 'end' in x else 'start')
logs_transformed['event_name'] = logs_transformed['event'].str.replace('_begins|_complete', '', regex=True)
logs_transformed['resource_id_full'] = logs_transformed.apply(lambda x: f"{x['event_name']}_{x['resource_id']:.0f}", axis=1)
logs_transformed = logs_transformed.sort_values(['run', 'time'], ascending=True)
# logs_transformed["activity_id"] = (
#     logs_transformed.groupby(["run", "patient", "event_name"]).ngroup() + 1
# )

# logs_transformed = logs_transformed.sort_values(["run", "patient", "activity_id", "event_stage"], ascending=[True, True, True, False])

# Sort the data by run, patient, time, and event_name to handle tied start times
logs_transformed = logs_transformed.sort_values(["run", "patient", "time", "event_name"])

# Get the first occurrence of each activity (the start event)
first_occurrences = (
    logs_transformed[logs_transformed["event_stage"] == "start"]
    .drop_duplicates(["run", "patient", "event_name"])
    .copy()
)

# Sort by time within each run to determine the proper sequence
first_occurrences = first_occurrences.sort_values(["run", "time", "event_name"])

# Assign sequential activity_id within each run
first_occurrences["activity_id"] = first_occurrences.groupby("run").cumcount() + 1

# Merge the activity_id back to the main DataFrame
logs_transformed = logs_transformed.merge(
    first_occurrences[["run", "patient", "event_name", "activity_id"]],
    on=["run", "patient", "event_name"],
    how="left"
)

# Sort for final ordering
logs_transformed = logs_transformed.sort_values(
    ["run", "patient", "activity_id", "event_stage"],
    ascending=[True, True, True, False]
)
logs_transformed.head(50)
```

```{python}
logs_transformed[logs_transformed["activity_id"]==26].sort_values('run').head(30)
```

```{python}
logs_transformed.sort_values('activity_id').head(20)
```

```{python}
logs_transformed[["event_name", "event_stage", "event_type"]].value_counts()
```

```{python}
logs_transformed.event.value_counts()
```

For ease, now let's save these results as a file that we can load into R.

We could use a csv for easy interoperability. Alternatively, we could use something like Feather or Parquet, which are usable by both R and Python while retaining data types.

For ease of use and long-term readbility, we will use csv in this case.

```{python}
logs_transformed.to_csv('simulation_logs_for_bupar.csv', index=False)
```

## Moving to R - Exploring the bupaR package

Now, it's time to move to R (as bupaR and the bupaverse is only implemented in R).

pm4py exists as a process analytics package for Python, but the visuals of bupaR are of a high quality.

### Importing the required R functions and our data

```{r}
library(dplyr)
library(readr)
library(bupaverse)
library(processanimateR)
library(lubridate)
library(DT)
library(psmineR)

data <- readr::read_csv("simulation_logs_for_bupar.csv")

data %>% head()
```

### Ensuring our data has the required time columns and is only for a single run

```{r}
#| label: preprocess-data

data_processed <- data %>%
    dplyr::filter(run == 0) %>%
    dplyr::rename(minutes_after_origin=time) %>%
    # We provide a theoretical date to act as a starting point - the date does not have to be
    # a true representation of the actual simulation, though you may wish it to be if there
    # are date elements in your simulation (e.g. within-week or within-year seasonality)
    dplyr::mutate(time = as.POSIXct("2024-01-01 00:00:00", tz = "UTC") + lubridate::dminutes(minutes_after_origin)) %>%
    bupaR::convert_timestamps("time", ymd_hms) %>%
    dplyr::mutate(patient = as.factor(patient))

DT::datatable(data_processed)
```

### Converting to the activitylog format

```{r}
#| label: make-activity-log

activity_log <- data_processed  %>%
    bupaR::eventlog(
        case_id = "patient",
        activity_id = "event_name",
        activity_instance_id = "activity_id",
        lifecycle_id = "event_stage",
        timestamp = "time",
        resource_id = "resource_id_full"
        )

## !!!! Note that the bupaR documentation recommmends using the
## to_activitylog() function at the end of this set of steps.
## This caused significant errors in testing of this code, so
## I would not recommend following this recommendation, and instead
## you can mimic the above

activity_log
```

## Creating outputs

### Process Maps

#### Absolute frequencies

```{r}
activity_log %>%
    process_map(frequency("absolute"))
```

```{r}
activity_log %>%
    process_map(frequency("absolute-case"))
```

#### Relative frequencies

```{r}
activity_log %>%
    process_map(frequency("relative"))
```

```{r}
activity_log %>%
    process_map(frequency("relative-case"),
                render_options = list(edge_label_color = "white"))
```

```{r}
activity_log %>%
    process_map(frequency("relative-consequent"),
                render_options = list(edge_label_color = "white"))
```

### Performance Maps

#### Mean Waits

```{r}
activity_log %>%
    process_map(performance())
```

#### Max Waits

```{r}
activity_log %>%
    process_map(performance(FUN = max))
```

#### 90th percentile

```{r}
p90 <- function(x, ...) {
    quantile(x, probs = 0.9, ...)
}

activity_log %>%
    process_map(performance(FUN = p90))
```

### Analytics

Take a look at [this page in the bupaR docs](https://bupaverse.github.io/docs/performance_analysis.html) details of each of these plots.

#### Idle Time

```{r}
activity_log %>%
    idle_time("resource", units = "mins")
```

```{r}
activity_log %>%
    idle_time("resource", units = "mins") %>%
    plot()
```

#### Processing Time

```{r}
activity_log %>%
    processing_time("log", units = "mins") %>%
    plot()
```

```{r}
activity_log %>%
    processing_time("case", units = "mins") %>%
    plot()
```

```{r}
activity_log %>%
    processing_time("activity", units = "mins") %>%
    plot()
```

```{r}
activity_log %>%
    processing_time("resource-activity", units = "mins") %>%
    plot()
```

#### Throughput time

```{r}
activity_log %>%
    throughput_time("log", units = "mins") %>%
    plot()
```

#### Activity Presence

```{r}
activity_log %>%
    activity_presence() %>%
    plot()
```


### Resource visualisations

#### Handover-of-work network

```{r}
activity_log %>%
    resource_map()
```

#### Resource precedence matrix

```{r}
activity_log %>%
    resource_matrix() %>%
    plot()
```

### Process matrix

```{r}
activity_log %>%
    process_matrix(frequency("absolute")) %>%
    plot()
```

### Trace Explorer

This plot helps us to unerstand how often different combinations of activities occur, and whether there are any unexpected paths in our data.

```{r}
activity_log %>%
    trace_explorer(n_traces = 10)
```

### Animated process map

```{r}
activity_log %>%
    animate_process()
```


Let's compare directly with our vidigi output.

:::{.callout-tip}
The key difference between what is produced via bupaverse's `animate_process` and what can be created via vidigi is the ability of vidigi to more clearly show the scale of queues, and the number of resources available at any given point.

Vidigi can also more clearly highlight the impact of priority on resources through the use of distinct icons, though this is not demonstrated in this example.
:::

:::{.callout-note collapse="true"}
### Click here to expand vidigi animation generation code

```{python}
#| label: rerun-run-python-model
#| eval: true
#| echo: true

from examples.example_2_branching_multistep.ex_2_model_classes import Trial, g
from vidigi.animation import animate_activity_log
import pandas as pd
import plotly.io as pio
#pio.renderers.default = "notebook"
pio.renderers.default = "iframe"

g.sim_duration = 3000
g.number_of_runs = 3

my_trial = Trial()

my_trial.run_trial()

event_position_df = pd.DataFrame([
    # {'event': 'arrival', 'x':  10, 'y': 250, 'label': "Arrival" },

    # Triage - minor and trauma
    {'event': 'triage_wait_begins',
        'x':  160, 'y': 375, 'label': "Waiting for<br>Triage"  },
    {'event': 'triage_begins',
        'x':  160, 'y': 315, 'resource':'n_triage', 'label': "Being Triaged" },

    # Minors (non-trauma) pathway
    {'event': 'MINORS_registration_wait_begins',
        'x':  300, 'y': 145, 'label': "Waiting for<br>Registration"  },
    {'event': 'MINORS_registration_begins',
        'x':  300, 'y': 85, 'resource':'n_reg', 'label':'Being<br>Registered'  },

    {'event': 'MINORS_examination_wait_begins',
        'x':  465, 'y': 145, 'label': "Waiting for<br>Examination"  },
    {'event': 'MINORS_examination_begins',
        'x':  465, 'y': 85, 'resource':'n_exam', 'label': "Being<br>Examined" },

    {'event': 'MINORS_treatment_wait_begins',
        'x':  630, 'y': 145, 'label': "Waiting for<br>Treatment"  },
    {'event': 'MINORS_treatment_begins',
        'x':  630, 'y': 85, 'resource':'n_cubicles_non_trauma_treat', 'label': "Being<br>Treated" },

    # Trauma pathway
    {'event': 'TRAUMA_stabilisation_wait_begins',
        'x': 300, 'y': 560, 'label': "Waiting for<br>Stabilisation" },
    {'event': 'TRAUMA_stabilisation_begins',
        'x': 300, 'y': 490, 'resource':'n_trauma', 'label': "Being<br>Stabilised" },

    {'event': 'TRAUMA_treatment_wait_begins',
        'x': 630, 'y': 560, 'label': "Waiting for<br>Treatment" },
    {'event': 'TRAUMA_treatment_begins',
        'x': 630, 'y': 490, 'resource':'n_cubicles_trauma_treat', 'label': "Being<br>Treated" },

    {'event': 'depart',
    'x':  670, 'y': 330, 'label': "Exit"}
    ])
```

```{python}
#| eval: false
#| echo: true
#| label: make-vidigi-anim-for-comparison

animate_activity_log(
        event_log=my_trial.all_event_logs[my_trial.all_event_logs['run']==0],
        event_position_df=event_position_df,
        scenario=g(),
        entity_col_name="patient",
        debug_mode=False,
        setup_mode=False,
        every_x_time_units=5,
        include_play_button=True,
        gap_between_entities=10,
        gap_between_resources=20,
        gap_between_queue_rows=20,
        plotly_height=600,
        plotly_width=1000,
        override_x_max=700,
        override_y_max=675,
        entity_icon_size=12,
        resource_icon_size=15,
        wrap_queues_at=10,
        step_snapshot_max=50,
        limit_duration=3000,
        time_display_units="dhm",
        display_stage_labels=False,
        add_background_image="https://raw.githubusercontent.com/Bergam0t/vidigi/refs/heads/main/examples/example_2_branching_multistep/Full%20Model%20Background%20Image%20-%20Horizontal%20Layout.drawio.png",
    )

```

:::

```{python}
#| echo: false
#| eval: true

animate_activity_log(
        event_log=my_trial.all_event_logs[my_trial.all_event_logs['run']==1],
        event_position_df= event_position_df,
        scenario=g(),
        entity_col_name="patient",
        debug_mode=False,
        setup_mode=False,
        every_x_time_units=5,
        include_play_button=True,
        gap_between_entities=10,
        gap_between_resources=20,
        gap_between_queue_rows=20,
        plotly_height=600,
        plotly_width=1000,
        override_x_max=700,
        override_y_max=675,
        entity_icon_size=12,
        resource_icon_size=15,
        wrap_queues_at=10,
        step_snapshot_max=50,
        limit_duration=3000,
        time_display_units="dhm",
        display_stage_labels=False,
        add_background_image="https://raw.githubusercontent.com/Bergam0t/vidigi/refs/heads/main/examples/example_2_branching_multistep/Full%20Model%20Background%20Image%20-%20Horizontal%20Layout.drawio.png",
    )

```

### Other chart types

#### Dotted chart

We can see the impact of the pattern of daily arrivals across the course of the model, with the waits clearing out overnight when arrivals slow down.

```{r}
activity_log %>%
    dotted_chart(x = "absolute")
```

```{r}
activity_log %>%
    dotted_chart(x = "relative", sort="start")
```


#### Breaking down dotted charts by route

##### Minors

```{r}
#| label: dot-minors-absolute
activity_log %>%
    filter(event_name %in% c('MINORS_examination', 'MINORS_registration', 'MINORS_treatment', 'triage')) %>%
    dotted_chart(x = "absolute")
```

```{r}
#| label: dot-minors-relative
activity_log %>%
    filter(event_name %in% c('MINORS_examination', 'MINORS_registration', 'MINORS_treatment', 'triage')) %>%
    dotted_chart(x = "relative", sort="start")
```

##### Trauma

```{r}
#| label: dot-trauma-absolute
activity_log %>%
    filter(event_name %in% c('TRAUMA_stabilisation', 'TRAUMA_treatment', 'triage')) %>%
    dotted_chart(x = "absolute")
```

```{r}
#| label: dot-trauma-relative
activity_log %>%
    filter(event_name %in% c('TRAUMA_stabilisation', 'TRAUMA_treatment', 'triage')) %>%
    dotted_chart(x = "relative", sort="start")
```

## Conclusion

vidigi and bupaR are complementary packages to use when visualising, verifying and validating your simulation models - or working with real-world process data.
